\textbf{Discrete Bulk Reconstruction Problem} \\ \large Compiled Research

### 2025-03-07 paper.tex
photos
added text

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{hyperref}
\usepackage{enumitem}

\title{DBRP Research Paper}
\author{}
\date{}

@@

\section{Task 1 - DBRP}

\subsection*{Goal}
\begin{itemize}
    \item Gain understanding on bulk reconstruction algorithm
    \item Source: Discrete bulk reconstruction, Scott Aaronson and Jason Pollack
\end{itemize}

\section{Background}

\subsection{AdS (Anti-de Sitter)}
\begin{itemize}
    \item Higher-dimensional \textbf{bulk} space
    \item \textbf{Bulk}: AdS space
    \item \textbf{Surface}: Geometric object within AdS space
    \item \textbf{Minimal area}: Surface that spans the boundary subregion and has the smallest possible area \textit{(RT surface in AdS/CFT)}
    \item The bulk is very hard to calculate directly because of quantum gravity, therefore perform calculations on CFT.
\end{itemize}

\subsection{CFT (Conformal Field Theory)}
\begin{itemize}
    \item \textbf{Boundary} $\Rightarrow$ CFT
    \item Lower-dimensional space that forms "edge" or "boundary" of AdS bulk
    \item Non-gravitational theory that lives on the boundary of the AdS spacetime
    \item Information of boundary is encoded in the bulk
\end{itemize}

@@

\subsection{AdS/CFT}
\begin{itemize}
    \item \textbf{AdS/CFT correspondence} $\Rightarrow$ The gravitational theory in the bulk is equivalent to the non-gravitational CFT on the boundary
    \item i.e. Two theories describe the same physics, but from different perspectives
    \item \textbf{Holographic Relationship}: Everything happening in the bulk AdS spacetime can be described by the boundary CFT, and vice versa
    \item Consists of two theories (theory of quantum gravity and a quantum field theory, with no gravity). The relation between the two theories is called "holographic".
    \item Theories are "equivalent"? (yet to be proven)
    \item Meaning, there exists a mapping or "dictionary" that maps all states and observables from one theory to corresponding states and observables in the other
    \item Currently lack a full dictionary for the AdS side
\end{itemize}

@@

\subsubsection{Spacetime Geometry for AdS/CFT}
\begin{itemize}
    \item In AdS space, geometry is negatively curved.
    \item \textbf{CFT can reconstruct the geometry of AdS bulk}: Meaning, data in the CFT can be used to determine how spacetime is shaped or curved in the AdS bulk.
    \item \textbf{Entanglement is directly related to geometry of the bulk spacetime}: 
    \begin{itemize}
        \item Amount of entanglement in different regions of CFT tells us about structure and curvature of the corresponding bulk spacetime in AdS.
        \item More entanglement on the boundary implies more curvature or structure in the bulk geometry.
        \item Proven with RT formula: more entanglement = larger minimal surface area = more curved and intricate bulk geometry.
    \end{itemize}
\end{itemize}

@@

\subsection{Entanglement Measurements}
\begin{itemize}
    \item Not needed for DBRP because problem states initially given full AdS/CFT correspondence.
\end{itemize}

@@

\section{Discrete Bulk Reconstruction Problem}
Official statement initialization:
\[
\mathcal{H} = \otimes^{N}_{i=1}\mathcal{H}_{i}
\]
\begin{quote}
    Note about tensor product, Hilbert space, and qubits:
\end{quote}
\begin{itemize}
    \item Tensor Product is a mathematical operation used to describe a system with multiple qubits. It combines two or more Hilbert spaces into a single composite Hilbert space, or one large vector space. The larger space captures the full state space of the combined quantum system.
    \item For a single qubit: $|0\rangle$, $|1\rangle$ form a basis for a two-dimensional Hilbert space.
    \item Therefore, an $n$-qubit system has a Hilbert space of dimension $2^n$.
\end{itemize}

@@

\section{Problem Set-Up}
Can model the bulk by a weighted undirected graph where each subregion is an RT (or minimal area) surface, which corresponds to min-cuts in \( G \).

\subsection{Official Statement}
Given as input a list of atomic boundary regions labeled \( 1, \dots, N \), a list of subsets of the regions \( R_1, \dots, R_k \subseteq [N] \), and a real-valued entropy \( S(R_i) \geq 0 \) for each \( R_i \), the weight of the minimum cut separating \( R_i \) from the rest of the boundary vertices (i.e., from \( [N] - R_i \)) is equal to \( S(R_i) \).

\subsection{General Idea}
\begin{itemize}
    \item If we are given the complete entanglement structure of the boundary, we are therefore provided with a min-cut graph of the boundary entropies. We can then perform an experiment in the bulk space and use min-cut theory on the boundary to reconstruct the entropy graph.
\end{itemize}

\textbf{This is basically the inverse min-cut problem.} Given the min-cut, find the graph that gave you the min-cut solution.

\section{Required Properties}
To ensure a solution exists we \textbf{need} the following properties:

\section{ ADD TWO PHOTOS}
\begin{enumerate}
    \item \textbf{Subadditivity (SA)}: entropy of \( A \) plus entropy of \( B \) is \( \geq \) entropy of \( A \cup B \)
    \[
    S(A) + S(B) \geq S(AB)
    \]
 %   \begin{center}
 %       \includegraphics[width=0.8\textwidth]{https://hackmd.io/_uploads/S1S-H-SzJe.png}
  %  \end{center}

    \item \textbf{Strong Subadditivity (SSA)}
    \[
    S(AB) + S(BC) \geq S(B) + S(ABC)
    \]
   % \begin{center}
  %      \includegraphics[width=0.8\textwidth]{https://hackmd.io/_uploads/S1vBSWBfyg.png}
  %  \end{center}
\end{enumerate}

\section{Holographic Quantum State}
\begin{itemize}
    \item Boundary state in CFT that, due to holography, has a corresponding interpretation in bulk AdS space.
    \item This allows us to perform computations on a lower-dimensional boundary with no gravity, which then gives us information about higher-dimensional gravity in bulk space.
    \item Note: Entanglement entropy of a region in boundary can correspond to the area of the minimal surface in the bulk (RT-formula).
\end{itemize}

All holographic quantum states \textbf{must} satisfy the following property:
\begin{itemize}
    \item \textbf{Monogamy of Mutual Information (MMI)}
    \[
    S(AB) + S(BC) + S(AC) \geq S(A) + S(B) + S(C) + S(ABC)
    \]
\end{itemize}

@@

\section{DBRP on 1-dimensional Hilbert Space}
\begin{itemize}
    \item Note: To perform DBRP on a higher-dimensional Hilbert space, take the tensor product of each Hilbert space to create one large Hilbert space.
    \[
    \mathcal{H}_{\text{total}} = \mathcal{H}_1 \otimes \mathcal{H}_2 \otimes \dots \otimes \mathcal{H}_N
    \]
    \item Upper bound for contiguous boundary regions: 
    \[
    \frac{N(N-1)}{2}
    \]
    \item Can specify \textit{all} \( 2^N \) non-contiguous entropies in terms of \( \frac{N(N-1)}{2} \) contiguous entropies.
    \item Formula is minimum of 2 different possibilities:
    \[
    S(AC) = \min\{ S(A) + S(C), S(B) + S(D) \}
    \]
 %%   \begin{center}
  %      \includegraphics[width=0.8\textwidth]{https://hackmd.io/_uploads/HkB0NMHG1e.jpg}
  %  \end{center}
\end{itemize}
\section{ADD PHOTO}

\section{Theorem: 1D Boundary}
In special cases of a 1D boundary divided into \( N \) parts, given contiguous entropies satisfying SSA, we can always find a graph model of the bulk in linear time \( O(N^2) \). The graph model is planar, universal, and has \textbf{only} \( O(N^2) \) vertices.

@@

\section{Proof of 1D, with Examples}
\begin{itemize}
    \item Any contiguous boundary data that satisfies SSA admits a "bulkless" graph.
    \item A bulkless graph is a graph of pure edges and no vertices.
    \item After creating a bulkless graph, we can solve for what the edge weights should be! By using a system of linear equations, SSA ensures nonnegativity.
\end{itemize}

\section{ADD PHOTO}
%\begin{center}
%    \includegraphics[width=0.8\textwidth]{https://hackmd.io/_uploads/H1d-6bSM1x.jpg}
%\end{center}

@@

\section{Chord Construction}
\begin{itemize}
    \item Get model of graph.
    \item Take all edges in between atomic boundary regions and add a chord (vertex) where each edge intersects.
    \item Edge weights = weights from bulkless graph.
\end{itemize}

\section{ADD PHOTO}
%\begin{center}
%    \includegraphics[width=0.8\textwidth]{https://hackmd.io/_uploads/HJ_pezHzyg.jpg}
%\end{center}

@@

\section{Diamondwork Construction}
\begin{itemize}
    \item Yields \( O(N^2) \) vertices and edges.
    \item Reduce via a planar graph made up of "overlapping" triangles, each contributing weight from the bulkless graph.
    \item Overlapping triangles are minimal surfaces that cut deeper into the bulk.
    \item Add edge weights together to overlap them.
    \item Min-cuts cut the same as the bulkless graph.
\end{itemize}

\section{ADD PHOTO}
%\begin{center}
%    \includegraphics[width=0.8\textwidth]{https://hackmd.io/_uploads/ry97mMHMkl.jpg}
%\end{center}

\section{Harmonic Edge Weights}
This is the discrete version of AdS geometry. The entropy of a length \( L \) boundary region (where \( L \leq N/2 \)) is given as:

\section{ADD PHOTO}
%\begin{center}
%    \includegraphics[width=0.8\textwidth]{https://hackmd.io/_uploads/B1S_4MBMJl.jpg}
%\end{center}

\section{Why No Black Holes?}
\begin{itemize}
    \item All edge weights have the same value, therefore the RT minimal surface means nothing.
    \item There is no advantage to cutting deeper into the bulk.
\end{itemize}

\section{ADD PHOTO}
%\begin{center}
%    \includegraphics[width=0.8\textwidth]{https://hackmd.io/_uploads/BJLSNo1VJe.jpg}
%\end{center}

>>>>

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{abstract}
\usepackage{geometry}

% page settings
\geometry{a4paper, margin=1in} % sets A4 paper with 1-inch margins

% Title Page
\title{\textbf{Discrete Bulk Reconstruction Problem} \\ \large Compiled Research}
\author{Piper Jeffries \\ Missouri S\&T \\ \texttt{pejbvy@umsystem.edu}}
\date{\today}

@@

\begin{abstract}
    This is example abstract. 
\end{abstract}

\newpage 
\tableofcontents
\newpage

\section{Introduction}
    Intro provides background info, motivation, and objective of research. Explain why problem is important and summarize prior work

\section{Methodology}
Describe your approach, methods, and any theoretical background required. If you include equations, use:

\section*{Background}

\subsection*{AdS (Anti-de Sitter)}
\begin{itemize}
    \item Higher-dimensional \textbf{bulk} space
    \item \textbf{Bulk}: AdS space
    \item \textbf{Surface}: Geometric object within AdS space
    \item \textbf{Minimal area}: Surface that spans the boundary subregion and has the smallest possible area \textit{(RT surface in AdS/CFT)}
    \item The bulk is very hard to calculate directly because of quantum gravity, therefore perform calculations on CFT.
\end{itemize}

\subsection*{CFT (Conformal Field Theory)}
\begin{itemize}
    \item \textbf{Boundary} $\Rightarrow$ CFT
    \item Lower-dimensional space that forms "edge" or "boundary" of AdS bulk
    \item Non-gravitational theory that lives on the boundary of the AdS spacetime
    \item Information of boundary is encoded in the bulk
\end{itemize}

@@

\subsection*{AdS/CFT}
\begin{itemize}
    \item \textbf{AdS/CFT correspondence} $\Rightarrow$ The gravitational theory in the bulk is equivalent to the non-gravitational CFT on the boundary
    \item i.e. Two theories describe the same physics, but from different perspectives
    \item \textbf{Holographic Relationship}: Everything happening in the bulk AdS spacetime can be described by the boundary CFT, and vice versa
    \item Consists of two theories (theory of quantum gravity and a quantum field theory, with no gravity). The relation between the two theories is called "holographic".
    \item Theories are "equivalent"? (yet to be proven)
    \item Meaning, there exists a mapping or "dictionary" that maps all states and observables from one theory to corresponding states and observables in the other
    \item Currently lack a full dictionary for the AdS side
\end{itemize}

@@

\subsubsection*{Spacetime Geometry for AdS/CFT}
\begin{itemize}
    \item In AdS space, geometry is negatively curved.
    \item \textbf{CFT can reconstruct the geometry of AdS bulk}: Meaning, data in the CFT can be used to determine how spacetime is shaped or curved in the AdS bulk.
    \item \textbf{Entanglement is directly related to geometry of the bulk spacetime}: 
    \begin{itemize}
        \item Amount of entanglement in different regions of CFT tells us about structure and curvature of the corresponding bulk spacetime in AdS.
        \item More entanglement on the boundary implies more curvature or structure in the bulk geometry.
        \item Proven with RT formula: more entanglement = larger minimal surface area = more curved and intricate bulk geometry.
    \end{itemize}
\end{itemize}

@@

\subsection*{Entanglement Measurements}
\begin{itemize}
    \item Not needed for DBRP because problem states initially given full AdS/CFT correspondence.
\end{itemize}

@@

\section*{Discrete Bulk Reconstruction Problem}
Official statement initialization:
\[
\mathcal{H} = \otimes^{N}_{i=1}\mathcal{H}_{i}
\]
\begin{quote}
    Note about tensor product, Hilbert space, and qubits:
\end{quote}
\begin{itemize}
    \item Tensor Product is a mathematical operation used to describe a system with multiple qubits. It combines two or more Hilbert spaces into a single composite Hilbert space, or one large vector space. The larger space captures the full state space of the combined quantum system.
    \item For a single qubit: $|0\rangle$, $|1\rangle$ form a basis for a two-dimensional Hilbert space.
    \item Therefore, an $n$-qubit system has a Hilbert space of dimension $2^n$.
\end{itemize}

@@

\section*{Problem Set-Up}
Can model the bulk by a weighted undirected graph where each subregion is an RT (or minimal area) surface, which corresponds to min-cuts in \( G \).

\begin{center}
    \includegraphics[width=0.8\textwidth]{min_cut_graph.pdf}
\end{center}

\subsection*{Official Statement}
Given as input a list of atomic boundary regions labeled \( 1, \dots, N \), a list of subsets of the regions \( R_1, \dots, R_k \subseteq [N] \), and a real-valued entropy \( S(R_i) \geq 0 \) for each \( R_i \), the weight of the minimum cut separating \( R_i \) from the rest of the boundary vertices (i.e., from \( [N] - R_i \)) is equal to \( S(R_i) \).

\subsection*{General Idea}
If we are given the complete entanglement structure of the boundary, we are therefore provided with a min-cut graph of the boundary entropies. We can then perform an experiment in the bulk space and use min-cut theory on the boundary to reconstruct the entropy graph. Giving us an interpretation of what the outcome would be in the bulk, because the boundary information is encoded in the bulk.

\textbf{This is basically the inverse min-cut problem.} Now we are given min-cut, find graph that gave you min-cut solution, therefore this graph is not unique. While the graph is not unique you can add on additional properties. For example, planar, few vertices. For this problem assume no wormholes. For DBRP to work we need to make ensure all quantum states satisfies properties (more on that later). \\
It is important to note and prove DBRP is computable, meaning it has an upper bound. \\
It has an upper bound of $2^{2^N}$

\section*{Proof:}
Given $K\leq2^N$ boundary regions form all $\leq2^N$ possible intersections of RT regions. At most 1 vertex needed in each region. Using linear programs for each edge weight takes "only" exp(exp(exp(N))) time!

\begin{center}
    \includegraphics[width=0.8\textwidth]{subregions.png}
\end{center}

\subsection*{How to determine edge weights?}
Use linear programming, for every possibility of what min-cuts could be try to solve some linear equations and equalities. Can we set edge weights so that it will be min-cut?
\\Therefore, gives us algo for deciding any instance of DBRP.

\section*{Required Properties}
To ensure a solution exists we \textbf{need} the following properties:

\begin{enumerate}
    \item \textbf{Subadditivity (SA)}: entropy of \( A \) plus entropy of \( B \) is \( \geq \) entropy of \( A \cup B \)
    \[
    S(A) + S(B) \geq S(AB)
    \]
    \begin{center}
        \includegraphics[width=\linewidth]{subaddivity.png}
    \end{center}

    \item \textbf{Strong Subadditivity (SSA)}
    \[
    S(AB) + S(BC) \geq S(B) + S(ABC)
    \]
    \begin{center}
        \includegraphics[width=0.8\textwidth]{strong_subaddivity.png}
    \end{center}
\end{enumerate}

\section*{Holographic Quantum State}
\begin{itemize}
    \item Boundary state in CFT that, due to holography, has a corresponding interpretation in bulk AdS space.
    \item This allows us to perform computations on a lower-dimensional boundary with no gravity, which then gives us information about higher-dimensional gravity in bulk space.
    \item Note: Entanglement entropy of a region in boundary can correspond to the area of the minimal surface in the bulk (RT-formula).
\end{itemize}

All holographic quantum states \textbf{must} satisfy the following property:
\begin{enumerate}
    \item \textbf{Monogamy of Mutual Information (MMI)}
    \[
    S(AB) + S(BC) + S(AC) \geq S(A) + S(B) + S(C) + S(ABC)
    \]
\end{enumerate}

@@

\section*{DBRP on 1-dimensional Hilbert Space}
\begin{itemize}
    \item Note: To perform DBRP on a higher-dimensional Hilbert space, take the tensor product of each Hilbert space to create one large Hilbert space.
    \[
    \mathcal{H}_{\text{total}} = \mathcal{H}_1 \otimes \mathcal{H}_2 \otimes \dots \otimes \mathcal{H}_N
    \]
\end{itemize}
\begin{enumerate}
    \item Upper bound for contiguous boundary regions: 
    \[
    \frac{N(N-1)}{2}
    \]
    \item Can specify \textit{all} \( 2^N \) non-contiguous entropies in terms of \( \frac{N(N-1)}{2} \) contiguous entropies.
    \item Formula is minimum of 2 different possibilities:
    \[
    S(AC) = \min\{ S(A) + S(C), S(B) + S(D) \}
    \]
    \begin{figure}[htbp]  % 'h' places the figure "here" (can be adjusted)
        \centering
        \includegraphics[width=0.8\textwidth]{parts.pdf}  % Adjust width as needed
        \caption{Where the first option is in red and second option is in blue}  % Caption text
        \label{fig:example}  % Label for referencing
    \end{figure}
\end{enumerate}

\section*{Theorem: 1D Boundary}
In special cases of a 1D boundary divided into \( N \) parts, given contiguous entropies satisfying SSA, we can always find a graph model of the bulk in linear time \( O(N^2) \). The graph model is planar, universal, and has \textbf{only} \( O(N^2) \) vertices.

@@

\section*{Proof of 1D, with Examples}
\begin{itemize}
    \item Any contiguous boundary data that satisfies SSA admits a "bulkless" graph.
    \item A bulkless graph is a graph of pure edges and no vertices.
    \item After creating a bulkless graph, we can solve for what the edge weights should be! By using a system of linear equations, SSA ensures nonnegativity.
\end{itemize}

\begin{center}
    \includegraphics[width=0.6\textwidth]{bulkless_graph.pdf}
\end{center}

@@

\section*{Chord Construction}
\begin{itemize}
    \item Get model of graph.
    \item Take all edges in between atomic boundary regions and add a chord (vertex) where each edge intersects.
    \item Edge weights = weights from bulkless graph.
\end{itemize}

\begin{center}
    \includegraphics[width=0.6\textwidth]{chord_construct.pdf}
\end{center}

@@

\section*{Diamondwork Construction}
\begin{itemize}
    \item Yields \( O(N^2) \) vertices and edges.
    \item Reduce via a planar graph made up of "overlapping" triangles, each contributing weight from the bulkless graph.
    \item Overlapping triangles are minimal surfaces that cut deeper into the bulk.
    \item Add edge weights together to overlap them.
    \item Min-cuts cut the same as the bulkless graph.
\end{itemize}

\begin{center}
    \includegraphics[width=0.6\textwidth]{edge_weights.pdf}
\end{center}

\section*{Harmonic Edge Weights}
This is the discrete version of AdS geometry. The entropy of a length \( L \) boundary region (where \( L \leq N/2 \)) is given as:

\begin{center}
    \includegraphics[width=0.8\textwidth]{harmonic.pdf}
\end{center}

\section*{Why No Black Holes?}
\begin{itemize}
    \item All edge weights have the same value, therefore the RT minimal surface means nothing.
    \item There is no advantage to cutting deeper into the bulk.
\end{itemize}

\begin{center}
    \includegraphics[width=0.8\textwidth]{dbrp_summed.pdf}
\end{center}

\section{Conclusion}
Summarize key findings, limitations, and possible future research directions

\newpage

%%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{9}
    \bibitem{example}
    Scott Aaronson and Jason Pollack, \textit{Discrete bulk reconstruction}, 2023.
    \bibitem{2}
    Ronak Ramachandran, \textit{Further Exploration of the Discrete Bulk Reconstruction Problem}, 2023.
    \bibitem{AI}
    ChatGPT
\end{thebibliography}
